Exploring probability distributions for bivariate temporal granularities

### 2019-08-12 main.Rmd
Introduction of linear time granularities

  The hierarchical structure of linear granularities creates a natural nested ordering. For example, hours are nested within days, days within weeks, weeks within months, and so on. We refer to granularities which are nested within multiple levels as "multiple-order-up" granularities. For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.

   Visualizing data across various granularities helps us to understand periodicities, pattern and anomalies in the data. Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach. This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in [@Wickham2009-pk]. 
bibliography: bibliography.bib
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    dev: "pdf"
---

@@

<!--temporal granularities and why should we care --> 
Temporal data can be available at various resolution depending on the context. Social and economic data are often collected and reported at coarses temporal scales like monthly, quarterly or annually. But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales than that was previously possible. For example, it might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimisation every second.  As the frequency of data increases, the number of questions about the observed variable that need to be addressed also increases. For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters. This approach requires deconstructing time in various possible ways.

@@

<!--existing closely related work and how ours relate --> 
[@Grolemund2011-qf] allows creation of granularities that are mostly single-order-up like hour of the day, second of the minute. This paper utilises the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.

@@

Time can be represented at varied levels of abstraction depending on the accuracy required for the context. Time granularity can be defined as the resolution power of the temporal qualification of a statement. Providing a formalism with the concept of time granularity is important to model time information across differently grained temporal domains [@Euzenat2005-de].

>>>>

  The hierarchical structure of linear granularities creates a natural nested ordering resulting in single-order-up and multiple-order-up granularities. For example, hour of the week and second of the hour are both multiple-order-up, while hour of the day and second of the minute are single-order-up.

   Visualizing data across granularities which are either single-order-up or multiple-order-up or periodic/aperiodic helps us to understand periodicities, pattern and anomalies in the data. Because of the large volume of data available, using displays of probability distributions conditional on one or more granularities is a potentially useful approach. This work provides tools for creating granularities and exploring the associated time series within the tidy workflow, so that probability distributions can be examined using the range of graphics available in [@Wickham2009-pk]. 
bibliography: bibliography.bib
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    dev: "pdf"
---

@@

<!--temporal granularities and why should we care --> 
Temporal data can be available at various resolution depending on the context. Social and economic data are often collected and reported at coarses temporal scales like monthly, quarterly or annually. But with recent advancement in technology, more and more data are recorded and stored at much finer temporal scales than that was previously possible. It might be sufficient to observe energy consumption every half an hour, but energy supply needs to be monitored every minute and number of web searches requires optimisation every second.  As the frequency of data increases, the number of questions about the observed variable that need to be addressed also increases. For example, data collected at an hourly scale can be analyzed using coarser temporal scales like days, months or quarters. This approach requires deconstructing time in various possible ways.

@@

<!--existing closely related work and how ours relate --> 
 This paper utilises the nestedness of time granularities to obtain multiple-order-up granularities from single-order-up ones.

@@

Often we partition time into years, months, weeks, days and so on in a hierarchical manner to relate it to data. Such discrete abstractions of time can be thought of as linear time granularities [@aigner2011visualization]. Providing a formalism to this abstraction is important to model a time series across differently grained temporal domains.
